### 2017-12-31 icde/revision/sec-exp-app.tex

\eat{
\begin{figure*}[tb!]
\vspace{-2ex}
\begin{minipage}[t]{0.5\textwidth}
\begin{center}
\subfigure[{\scriptsize TWPageRank (batch vs. inc.)}]{\label{fig-aminer-time2}
\includegraphics[scale=0.38]{./exp/AMiner_time_twpr.eps}}
%\quad\quad
\hspace{-0.5ex}
\subfigure[{\scriptsize Comparison of ranking algorithms}]{\label{fig-aminer-time1}
\includegraphics[scale=0.38]{./exp/AMiner_time.eps}}
\end{center}
\vspace{-5ex}
\caption{\small Efficiency tests on \aminer}
\label{fig-aminer-time}
\end{minipage}
\hspace{-2ex}
\begin{minipage}[t]{0.5\textwidth}
\begin{center}
\subfigure[{\scriptsize \aminer}]{\label{fig-aminer-time-sigma}
\includegraphics[scale=0.38]{./exp/AMiner_sigma_time.eps}}
%\quad\quad
\hspace{-0.5ex}
\subfigure[{\scriptsize \magdata}]{\label{fig-mag-time-sigma}
\includegraphics[scale=0.38]{./exp/MAG_sigma_time.eps}}
\end{center}
\vspace{-5ex}
\caption{\small Efficiency tests: varying $\sigma$}
\label{fig-time-sigma}
\end{minipage}
\vspace{-1ex}
\end{figure*}
}

\newcommand{\graphscaleexpapp}{0.262}
\newcommand{\graphmarginexpapp}{-2ex}
%%% all in 1 Figure
\begin{figure*}[tb!]
\addtolength{\subfigcapskip}{-1ex}
\begin{center}
\subfigure[{\scriptsize \aan}]{\label{fig-aan-ab-recom}
\includegraphics[scale=\graphscaleexpapp]{./exp/AAN-para-recm.eps}}
%\quad\quad
\hspace{\graphmarginexpapp}
\subfigure[{\scriptsize \aminer}]{\label{fig-aminer-ab-recom}
\includegraphics[scale=\graphscaleexpapp]{./exp/AMiner-para-recm.eps}}
%\quad\quad
\hspace{\graphmarginexpapp}
\subfigure[{\scriptsize \magdata}]{\label{fig-mag-ab-recom}
\includegraphics[scale=\graphscaleexpapp]{./exp/MAG-para-recm.eps}}
\\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-2ex}
%\hspace{-10ex}
\subfigure[{\scriptsize \aan}]{\label{fig-aan-ab-fcita}
\includegraphics[scale=\graphscaleexpapp]{./exp/AAN-para-fcita.eps}}
%\quad\quad
\hspace{\graphmarginexpapp}
\subfigure[{\scriptsize \aminer}]{\label{fig-aminer-ab-fcita}
\includegraphics[scale=\graphscaleexpapp]{./exp/AMiner-para-fcita.eps}}
%\quad\quad
\hspace{\graphmarginexpapp}
\subfigure[{\scriptsize \magdata}]{\label{fig-mag-ab-fcita}
\includegraphics[scale=\graphscaleexpapp]{./exp/MAG-para-fcita.eps}}
\end{center}
\vspace{-3.5ex}
\caption{\small Accuracy tests with \recom ((a)--(c)) and \fcita ((d)--(f)): varying aggregating parameters $\alpha$ and $\beta$}
\label{fig-ab}
\vspace{-3ex}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tb!]
%\vspace{-2ex}
\begin{center}
\begin{small}
\vspace{1ex}
\begin{tabular}{|c|c c c|c c c| c|}
\hline
{\bf Datasets} & {\bf C} & {\bf V} & {\bf A} & {\bf CV} & {\bf CA} & {\bf VA} & {\bf CVA} \\
\hline \hline
% \recom
\aan & 0.746 & 0.616 & 0.647 & 0.806 & 0.760 & 0.746 & {\bf \ 0.810\ } \\
\aminer & 0.783 & 0.576 & 0.664 & 0.805 & 0.798 & 0.735 & {\bf 0.815} \\
\magdata & 0.669 & 0.534 & 0.552 & 0.693 & 0.670 & 0.640 & {\bf 0.694} \\ \hline
% \ficta
\aan & 0.708 & 0.566 & 0.744 & 0.778 & 0.796 & 0.755 & {\bf 0.804} \\
\aminer & 0.690 & 0.733 & 0.761 & 0.854 & 0.839 & 0.794 & {\bf 0.857} \\
\magdata & 0.761 & 0.657 & 0.737 & 0.835 & 0.842 & 0.753 & {\bf 0.849} \\
\hline
\end{tabular}
\end{small}
\end{center}
\caption{\small Accuracy tests on using different components with \recom (rows 2--4) and \fcita (rows 5--7), in which letters C, V and A stand for citation, venue and author components, respectively.}
\label{tab-inc}
\vspace{-6ex}
\end{table}
%%%%%%%%%%%%%%%%%%%

\vspace{-1ex}
\section*{APPENDIX A: Detailed results of Exp-4.3}
\label{sec-exp-app}

%\subsection*{1. Detailed results of Exp-4.3} %$\alpha$ and $\beta$ (detailed)

%We present detailed results of {\em Exp-4.3}, which are omitted earlier due to the space constraints, \ie the impacts of aggregating parameters $\alpha$ and $\beta$ on effectiveness.

\eat{
\etitle{Exp-4.1}. To evaluate the impacts of the time decaying factor $\sigma$, we varied $\sigma$ from -1.6 to -0.4, while fixed $Y_s$ to default values, $b=+\infty$ and $dif=1$. The efficiency results are reported in Fig.~\ref{fig-time-sigma}.
Note that the running time of algorithm \batensemble is different on \recom and \fcita because of the different splitting year $Y_s$.

When varying $\sigma$, the running time of \batensemble is almost stable on both \aminer and \magdata using \fcita and \recom. Indeed, the running time using (\fcita, \recom) only varies (1.4\%, 0.5\%) on \aminer, and (2.1\%, 2.2\%) on \magdata, on average, respectively.
}

\etitle{Exp-4.3}.
To evaluate the impacts of aggregating parameters $\alpha$ and $\beta$, we varied $\alpha$ and $\beta$ at the granularity of 0.01.
The detailed results are reported in Fig.~\ref{fig-ab}, where the parameters selected earlier and their corresponding \PairAcc are marked with $*$.

When varying $\alpha$ and $\beta$, the \PairAcc of \ensemblerank changes gently, as shown in Fig.~\ref{fig-ab}.
The optimal \PairAcc is obtained within a single region, rather than a complex collection of optimal regions.
%
Moreover, the \PairAcc keeps at a high level within a certain ($\alpha$, $\beta$) combination space around the optimal region, as shown in Fig.~\ref{fig-ab}.
%For instance, consider a square of length 0.3, which covers 8.5\% of the parameter combination space. The fraction of parameters such that the \PairAcc is no worse than 1\% of the corresponding \PairAcc with marker $*$ is (73\%, 94\%) on \aan, (96\%, 87\%) on \aminer and (83\%, 95\%) on \magdata, using (\recom, \fcita), respectively.
%
Further, the optimal parameters on the same sets of ground-truth are very similar for (\aan, \aminer and \magdata), indicating that the setting of $\alpha$ and $\beta$ can be easily transferred across different datasets.
To conclude, \ensemblerank is very robust to parameters $\alpha$ and $\beta$, and it is quite flexible for choosing proper values of parameters $\alpha$ and $\beta$.

Moreover, varying $\alpha$ and $\beta$ also enables to verify the effectiveness of importance assembling from different components, whose results are reported in Table~4.
The ranking based on all components consistently performs the best, which, using \recom and \fcita, improves the \PairAcc over using components (C, V, A, CV, CA, VA) by (6.33\%, 19.4\%, 16.3\%, 0.33\%, 4.97\%, 6.33\%) and (9.58\%, 23.8\%, 5.98\%, 2.56\%, 0.74\%, 4.82\%) on \aan, (3.21\%, 23.9\%, 15.1\%, 0.93\%, 1.66\%, 7.97\%) and (16.8\%, 12.5\%, 9.60\%, 0.34\%, 1.87\%, 6.37\%) on \aminer, and ( 2.43\%, 16.0\%, 14.2\%, 0.05\%, 2.38\%, 5.38\%) and (8.89\%, 19.3\%, 11.2\%, 1.43\%, 0.77\%, 9.65\%) on \magdata, respectively.
%Here the letters C, V and A stand for citation component, venue components and author component, respectively.