### 2018-01-22 ApproxComp/sec-introduction.tex

\section{Introduction}
\label{sec-intro}

1. history  of big data - governmental policy.

%2012年3月29日，美国总统科技政策办公室OSTP（Office of Science and Technology Policy）宣布了每年投资两亿美元的“大数据研究计划”（Big Data R&D Initiative）
%同天，我国科技部发布的“‘十二五’国家科技计划信息技术领域2013年度备选项目征集指南”把“大数据研究”列在首位

%美国2016年5月发布《联邦大数据研究与开发战略计划》
%其目标是对联邦机构的大数据相关项目和投资进行指导,主要围绕代表大数据研发关键领域的七个战略进行，包括促进人类对科学、医学和安全所有分支的认识;确保美国在研发领域继续发挥领导作用；通过研发来提高美国和世界解决紧迫社会和环境问题的能力。

2. progress of applications of big data techniques

3. (machine learning, systems -hadoop, spark)

A fundamental issue for big data analytics is the efficiency. There are different recent advances towards attacking this issues, from theory to algorithms to platforms~\cite{FanGN13,Jordan15,ZahariaXWDADMRV16}.  In this article, we present the idea of two classes of techniques for efficient and effective big
data analytics: query approximation and data approximation, based on our recent research. Different from existing approximation techniques~\cite{CormenLRS01,approx03,FanH14}, the approximation techniques that we are going to introduce are not unnecessarily theoretically bounded with respect to optimal solutions, but are practically useful and interesting.

~\cite{MaLHLH16}

Full-length Contributed Articles should consist of up to 4,000 words, contain no more than 25 references, 3-4 tables, 3-4 figures, and be submitted to: http://cacm.acm.org/submissions.

https://cacm.acm.org/about-communications/author-center/author-guidelines