### 2018-01-25 ApproxComp/sec-beyond.tex

\section{Beyond Query and Data Approximations}
\label{sec-beyond}

  For big data analytics, there are no one-size-fits-all techniques.  it is often necessary to combine different techniques to obtain good solutions.
We have seen that data sampling helps to achieve a balance between efficiency and effectiveness for link prediction,
and other techniques, \eg incremental computation, data partitioning, data indexing and distributed computing~\cite{MaLHLH16} can
also be unitized to  design query and data approximation techniques for big data analytics. 
For instance, it is easy to see that incremental algorithms can be utilized to achieve query approximation.