### 2018-01-25 ApproxComp/sec-conclusion.tex

\section{Conclusion}
\label{sec-conclusion}

In this article we have systematiclly introduced approximation computation techniques toward efficient and effective big data analytics.
Furthermore, although approximate computation does not put
theoretical bounds with respect to optimal solutions, it does expect a balance between efficiency and accuracy. Indeed, (1) our query approximation techniques~\cite{tods-MaCFHW14,LinMZWH17,MaHWLH17} show that efficiency can be obtained efficiency with high accuracy in practice, and (2) our data approximation techniques for shortest paths and distances and link prediction~\cite{MaFLWCH16,DuanMAMH17} show that efficiency and accuracy can be obtained for certain data analytic tasks. That is, the design policy approximate computation of is not to sacrifice accuracy for efficiency, but to achieve both efficiency and effectiveness in practice.