### 2018-01-26 ApproxComp/sec-beyond.tex

\section{Beyond Query and Data Approximation}
\label{sec-beyond}

  For big data analytics, there are no one-size-fits-all techniques, and it is often necessary to combine different techniques to obtain good solutions.

We have seen that data sampling helps to achieve a balance between efficiency and effectiveness for link prediction,
and other techniques such as incremental computation, data partitioning, data indexing and distributed computing can
also be unitized for big data analytics, and even for designing query and data approximation techniques for big data analytics.
For instance, it is easy to see that incremental algorithms (see~\cite{Reps96} for a survey) can be utilized to achieve query approximation.

It is also obvious that techniques beyond query and data approximation, such as theory and systems~\cite{FanGN13,Jordan15,ZahariaXWDADMRV16}, are must things for big data analytics.