### 2018-01-27 ApproxComp/coverletter.tex

\documentclass{letter}
\usepackage{geometry}

% duan
\usepackage{xspace}

\geometry{left=2.0cm, right=2.0cm, top=2.5cm, bottom=2.5cm}
\newcommand{\ie}{\emph{i.e.,}\xspace}
\newcommand{\eg}{\emph{e.g.,}\xspace}
\newcommand{\wrt}{\emph{w.r.t.}\xspace}
\newcommand{\aka}{\emph{a.k.a.}\xspace}
\newcommand{\kwlog}{\emph{w.l.o.g.}\xspace}
\newcommand{\etal}{\emph{et al.}\xspace}
\newcommand{\sstab}{\rule{0pt}{8pt}\\[-2.4ex]}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}

\begin{document}

Dear Editors,

I would like to request you to consider the attached manuscript entitled ``Approximate Computation for Big Data Analytics'' for publication in Communications of the ACM  as a contributed article.

The central theme of the article is big data analytics. A fundamental issue for big data analytics is the efficiency, and in this article we systematically introduce approximate computation (query approximation and data approximation) for efficiency and effectiveness big data analytics in practice.
I believe that the findings of this study are relevant to the scope of your journal, and will be of interests to the corresponding readers.

No conflicts of interest exits in the submission of this manuscript, and this manuscript is approved by all authors for publication, and  all the authors listed have approved the manuscript that is enclosed.

Below please find the names and email addresses of possessible appropriate reviewers:

\bi
\item[(1)] Prof. Sihem Amer-Yahia: Sihem.Amer-Yahia@imag.fr

\item[(2)]  Prof. Evaggelia Pitoura: pitoura@cs.uoi.gr

\item[(3)]  Prof. Lei Chen: leichen@cse.ust.hk

\item[(4)]  Prof. Feifei Li: lifeifei@cs.utah.edu
\ei

Looking forward to hearing from you.

Your sincerely,

Shuai Ma

Email: mashuai@buaa.edu.cn\\
Homepage: http://mashuai.buaa.edu.cn/
\end{document}