### 2018-09-23 ApproxComp/TKDE/sec-conclusion.tex

\section{Conclusions}
\label{sec-conclusion}

In this article we have systematiclly introduced approximation computation techniques for efficient and effective big data analytics.
Furthermore, although approximate computation does not put
theoretical bounds with respect to optimal solutions, it does expect a balance between efficiency and effectiveness. Indeed, (a) our query approximation techniques~\cite{ShuaiMaVLDB12,tods-MaCFHW14,LinMZWH17,MaHWLH17} show that efficiency can be obtained with high accuracy in practice, and (b) our data approximation techniques for shortest paths/distances and link prediction~\cite{MaFLWCH16,MaFLWCH17,HuAMH16,DuanAMHH16,DuanMAMH17} show that efficiency and accuracy can be obtained simultaneously for certain data analytic tasks. That is, the design policy of approximate computation is not to sacrifice effectiveness for efficiency, but to achieve both efficiency and effectiveness in practice.