### 2018-09-23 ApproxComp/TKDE/sec-beyond.tex

\section{Beyond Approximation Techniques}
\label{sec-beyond}

For big data analytics, there are no one-size-fits-all techniques, and it is often necessary to combine different techniques to obtain good solutions.

We have seen that sampling helps to achieve a balance between efficiency and effectiveness for approximate query processing~\cite{ChaudhuriDK17,Mozafari17,Kraska17,GarofalakisG01} and link prediction~\cite{DuanAMHH16,DuanMAMH17}},
and other techniques such as incremental computation~\cite{Reps96,FanLMTWW10,rankicde2018}, distributed computing~\cite{MaCHW12,FanXWYJZZCT17}, and system techniques~\eg caching~\cite{WangLMNT18}, hardware~\cite{AbergerLTNOR17,Han0Y18} can also be unitized for big data analytics, and can even be combined for designing query and data approximation techniques for big data analytics.

It is worth pointing out that (1) for all kind of techniques big data analytics, various computing resources should be seriously considered, \eg using bounded resources for approximation~\cite{CaoF17} and for incremental computation~\cite{abs-1801-01012}, and (2) theoretical analyses are also important for developing approximation techniques. For instance, our query and data approximation techniques are based serious theoretical results~\cite{ShuaiMaVLDB12,tods-MaCFHW14,LinMZWH17,MaFLWCH16,MaFLWCH17}.